---
title: "bayes HR"
author: "Jiahao Tian"
date: "2023-05-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(readr)
library(tumgr)
library(minpack.lm)
library(survival)
library(survminer)
library(grid)
library(lme4)
library(rstanarm)
library(bayesplot)
library(loo)
```

````{r}

# Create a new binary column based on the existing column
t_df22$decay = ifelse(t_df22$group %in% c("low decay", "high decay"), 1, 0)
t_df22$growth = ifelse(t_df22$group %in% c("low growth", "high growth"), 1, 0)


```


```{r}
## fit cox ph model
cox_model = coxph(Surv(time_mon, Death) ~ group, data = t_df22, iter.max = 2000)

summary(cox_model)
```




```{R}

# low_decay + high_decay + low_growth + high_growth + stable + yoyo

# Define the prior
# Fit the model
library(rstanarm)
prior_fit = stan_surv(Surv(time_mon, Death) ~ group,
                data = t_df22, 
                basehaz = "exp",
                #prior = normal(autoscale = T),
                #prior_intercept = normal(autoscale = T),
                #prior_aux = cauchy(0, 5, autoscale = T),
                #prior_PD = T,
                cores = 4,
                chains = 3, 
                iter = 5000, 
                seed = 123)
```

```{r}
summary(mod1_weibull)
prior_summary(mod1_weibull)
```


########################################################################################

```{r}
#mod1_exp <- update(prior_fit, basehaz = "exp")
mod1_weibull <- update(prior_fit, basehaz = "weibull")
mod1_gompertz <- update(prior_fit, basehaz = "gompertz")
mod1_bspline <- update(prior_fit, basehaz = "bs")
mod1_mspline1 <- update(prior_fit, basehaz = "ms")
mod1_mspline2 <- update(prior_fit, basehaz = "ms", basehaz_ops = list(df = 9))

```

```{r}
loo_compare(loo(prior_fit),
            loo(mod1_weibull),
            loo(mod1_gompertz),
            loo(mod1_bspline),
            loo(mod1_mspline1),
            loo(mod1_mspline2))


```


```{r}
nd = data.frame(group = c("low decay",
                          "high decay",
                          "low growth",
                          "high growth",
                          "stable",
                          "yoyo",
                          "remainder"))
head(nd)

ps <- posterior_survfit(prior_fit,
                        newdata = nd,
                        times = 0,
                        extrapolate = TRUE,
                        control = list(edist = 5))
head(ps)
```

```{r}


panel_labels <- c('1' = "low decay",
                  '2' = "high decay",
                  '3' = "low growth",
                  '4' = "high growth",
                  '5' = "stable",
                  '6' =  "yoyo",
                  '7' = "remainder")

pps <- plot(ps) + facet_wrap(~ id, labeller = labeller(id = panel_labels))
```


```{r}
ph <- posterior_survfit(mod1_weibull, newdata = nd, type = "haz")
pl <- posterior_survfit(mod1_exp, newdata = nd, type = "loghaz")

pph <- plot(ph) +
  facet_wrap(~ id, labeller = labeller(id = panel_labels))
ppl <- plot(pl) +
  facet_wrap(~ id, labeller = labeller(id = panel_labels))

```


########################################################################################
```{R}
## inspect the marginal prior distributions for log HR for all rate groups
mcmc_intervals(prior_fit, pars = c("grouphigh growth",
                                          "grouplow decay",
                                          "grouplow growth",
                                          "groupremainder",
                                          "groupstable",
                                          "groupyoyo",
                                          "(Intercept)"))

mcmc_intervals(prior_fit, pars = c("grouphigh growth",
                                          "grouplow decay",
                                          "grouplow growth",
                                          "groupremainder",
                                          "groupstable",
                                          "groupyoyo"),
               transformations = exp) + vline_at(1)

###the plot of the marginal prior distribution for our HR shows very extreme values (no good)
```

```{R}
## inspect the marginal prior distributions for log HR for all rate groups
mcmc_intervals(prior_fit, pars = c("low_decay",
                                          "high_decay",
                                          "low_growth",
                                          "high_growth",
                                          "stable",
                                          "yoyo",
                                          "(Intercept)"))

mcmc_intervals(prior_fit, pars = c("low_decay",
                                          "high_decay",
                                          "low_growth",
                                          "high_growth",
                                          "stable",
                                          "yoyo"),
               transformations = exp) + vline_at(1)

###the plot of the marginal prior distribution for our HR shows very extreme values (no good)
```



```{R}
plot(prior_fit, plotfun = "basehaz") ##no good
plot(prior_fit, "dens_overlay")
plot(prior_fit, "trace")
```




```{R}
### upate prior
mcmc_intervals(prior_fit, pars = c("grouphigh growth",
                                          "grouplow decay",
                                          "grouplow growth",
                                          "groupremainder",
                                          "groupstable",
                                          "groupyoyo",
                                          "(Intercept)")) ### still no good


```

```{R}
### now conditional on our data
fit = update(prior_fit, prior_PD = FALSE)
print(fit, digits = 3)

## much better 

```


```{R}
library(cowplot)
## visualize the posterior results and compare to priors

bayesplot_grid(mcmc_intervals(prior_fit),
               mcmc_intervals(fit),
               titles = c("Prior","Posterior"))
          
bayesplot_grid(mcmc_hist(prior_fit),
               mcmc_hist(fit),
               titles = c("Prior","Posterior"),
               grid_args = list(nrow = 2))

## Here we see that our estimated posterior distributions are not at all incompatible with our prior beliefs. This is reassuring!!!!!

```

```{r}
## harzad ratio
## visually compare the prior and posterior for the estimated HR. 
## For comparison also show the HR calculated using the Cox proportional hazards model (vertical red line)

add_cox_hr = vline_at(exp(coef(cox_model)), color = "red")

bayesplot_grid(mcmc_hist(prior_fit,
                         pars = c("grouphigh growth",
                                          "grouplow decay",
                                          "grouplow growth",
                                          "groupremainder",
                                          "groupstable",
                                          "groupyoyo"),
                         transformations = exp,
                         binwidth = 0.05) + add_cox_hr,
               
               mcmc_hist(fit,
                         pars = c("grouphigh growth",
                                          "grouplow decay",
                                          "grouplow growth",
                                          "groupremainder",
                                          "groupstable",
                                          "groupyoyo"),
                         transformations = exp,
                         binwidth = 0.05) + add_cox_hr,
               titles = c("Prior","Posterior"),
               xlim = c(0,3),
               grid_args = list(nrow = 2))
```


```{r}
## harzad ratio
## visually compare the prior and posterior for the estimated HR. 
## For comparison also show the HR calculated using the Cox proportional hazards model (vertical red line)

add_cox_hr = vline_at(exp(coef(cox_model)), color = "red")

bayesplot_grid(mcmc_hist(prior_fit,
                         pars = c("low_decay",
                                          "high_decay",
                                          "low_growth",
                                          "high_growth",
                                          "stable",
                                          "yoyo"),
                         transformations = exp,
                         binwidth = 0.05) + add_cox_hr,
               
               mcmc_hist(fit,
                         pars = c("low_decay",
                                          "high_decay",
                                          "low_growth",
                                          "high_growth",
                                          "stable",
                                          "yoyo"),
                         transformations = exp,
                         binwidth = 0.05) + add_cox_hr,
               titles = c("Prior","Posterior"),
               xlim = c(0,3),
               grid_args = list(nrow = 2))
```


- Since this pattern is observed for all groups it might suggest that this is an issue with the estimated baseline hazard (i.e.(Intercept)) rather than the estimated effect of other groups. 
- The exponential proportional hazards model assumes a constant baseline hazard, which may be unrealistic for this data i believe.... 
- Need to consider a more flexible form for the baseline hazard, while still assuming proportional hazards for the effect of each rate groups.


```{r}
## model estimation

# weibull model
fit_weib = update(fit,
                  basehaz = "weibull")

# cubic m-spline model (with df = 5)
fit_ms5 = update(fit,
                 basehaz = "ms",
                 basehaz_ops = list(df = 5))

# cubic m-spline model (with df = 10)
fit_ms10 = update(fit,
                  basehaz = "ms",
                  basehaz_ops = list(df = 10))

# piecewise constant model (with df = 5)
fit_pw5 = update(fit,
                 basehaz = "ms",
                 basehaz_ops = list(degree = 0, df = 5))

# piecewise constant model (with df = 10)
fit_pw10 = update(fit,
                  basehaz = "ms",
                  basehaz_ops = list(degree = 0, df = 10))

fits_stan = list("Exp" = fit,
                 "Weibull" = fit_weib,
                 "MS (df = 5)" = fit_ms5,
                 "MS (df = 10)" = fit_ms10,
                 "PW (df = 5)" = fit_pw5,
                 "PW (df = 10)" = fit_pw10)

```



```{R}
plots = map(fits_stan, plot)

bayesplot_grid(plots = plots,
               ylim = c(0, 0.1),
               titles = names(fits_stan),
               grid_args = list(ncol = 3))

## It is apparent from the plot that the constant baseline hazard (i.e. exponential model) is inappropriate 
## it fails to capture any changes in the baseline hazard. 
## the other baseline hazards quite capture changes in the baseline hazard over time
```

```{R}
## For each model can compare the posterior estimate of the standardised survival curve to the Kaplan-Meier
## This helps to assess how well each model fits the observed data. 

# define helper function to add knot locations 
add_knots = function(x) {
  knots = x$basehaz$knots
  if (is.null(knots))
    return(NULL) 
  geom_vline(xintercept = knots, color = "green", alpha = 0.5)
}

# generate the 'ps_check' 
plots = map(fits_stan, ~ (ps_check(.) + add_knots(.)))

# combine the plots
bayesplot_grid(plots = plots,
               titles = names(fits_stan),
               grid_args = list(ncol = 3))

## Again we see evidence that the constant (i.e. exponential) model fits very poorly. 
## The Weibull model also appears no good. 
## The remaining models appear to do reasonably well, although there is some evidence that the M-spline model with only 5 degrees of freedom may not provide sufficient flexibility.

```


```{r}
## leave-one-out cross validation and modelcomparison based on the expected log predictive density

loos = map(fits_stan, loo)
loo_compare(loos)

```

```{r}
round(as.data.frame(summary(mod1_exp)), 2)


```

##############brm


```{r}
prior1 <- normal(location = 0, scale = 10, autoscale = T)
f1 = Surv(time_mon, Death) ~ tve(high_decay) + low_decay
 
fit <- stan_surv(
  f1,
  data = sub_data,
  basehaz = "exp",
  prior = prior1,
  chains = 4,
  iter = 2000)



```

```{r}

summary(fit)
exp(fit$coefficients)

plot(fit, "trace")


```




```{r}
# Assuming your original dataframe is named 'your_data' and you want to select columns 'column1', 'column3', 'column6', and 'column8'
sub_data <- t_df22[, c("time_mon", "Death", "name", "low_decay", "high_decay")]



```